---
title: "Naive T cell scRNA-seq import and analysis"
authors: "Dan Bunis, Rebecca Jaszczak"
date: "10/10/2022"
output:
  html_document:
    toc: true
    theme: united
---

# Installing one more package helpful for plotting figures.

```{r one more install, include=FALSE}
install.packages('gridExtra')
```

# Read in libraries and set notebook defaults.

```{r setup & libraries, include=FALSE}
#set some Rmd defaults
knitr::opts_chunk$set(echo = TRUE, fig.height = 6, fig.width = 7, warning = F)

#load in libraries
library(dittoSeq)
library(tidyverse)
library(Seurat)
library(ggplot2)
```

# Catching up on PCA, Neighbors Cal, Clustering, and UMAP, and saving

## Re-initialize Tcells data after our 'save point' from last session.

```{r read back in}
Tcells <- readRDS("../processed_data/Tcells_analyzed.rds")
```

Rebecca forgot to have us save the processed object after doing PCA, Find Neighbors, and FindClusters - oops! Let's rerun the code and do that now.

I wanted to spend just a few minutes on PC analysis, since we discussed it a bit in the previous two lectures, but maybe didn't cover it as thoroughly as I should have.

Next we perform Principle Component Analysis (PCA) on the scaled data. You can chose to run this on the 2000 Highly Variable Genes, or on all genes. `DimHeatmap()` allows for easy exploration of the primary sources of heterogeneity in a dataset, and can be useful when trying to decide which PCs to include for further downstream analyses.     

To overcome the extensive technical noise in any single feature for scRNA-seq data, Seurat clusters cells based on their PCA scores, with each PC essentially representing a ‘metafeature’ that combines information across a correlated feature (gene) set. The top principal components therefore represent a robust compression of the dataset. 

More info on PCA: https://hbctraining.github.io/scRNA-seq/lessons/05_normalization_and_PCA.html


```{r re-run PCA/clustering/umap}
Tcells <- RunPCA(
  object = Tcells,
  verbose = FALSE,
  npcs = 12)

pca_dims = 1:9

Tcells <- FindNeighbors(object = Tcells,
                        reduction = "pca",
                        k.param = 20,
                        dims = pca_dims)

Tcells <- FindClusters(object = Tcells,
                       modularity.fxn = 1,
                       algorithm = 1,
                       resolution = c(0.1,0.2))
Tcells <- RunUMAP(
  object = Tcells,
  reduction = "pca",
  dims = pca_dims,
  seed.use = 1)
```

```{r plot UMAP}
#Colored with CD4s lighter and CD8s darker 
dittoDimPlot(object = Tcells, 
             var =  "Tage",
             size = 1.5,
             reduction.use = "umap",
             colors = c(1:3,9:11),
             main = "T cells Lineage and Stage")

dittoDimPlot(object = Tcells, 
             var =  "RNA_snn_res.0.1",
             size = 1.5,
             reduction.use = "umap")

dittoDimPlot(object = Tcells, 
             var =  "RNA_snn_res.0.2",
             size = 1.5,
             reduction.use = "umap")
```

```{r save!}
# saveRDS(Tcells, "../processed_data/Tcells_analyzed.rds")
```


# Let's make the top of Figure 3!

```{r main umap, fig.height=3, fig.width=4}



# legend = dittoSeq:::.grab_legend(
#   dittoDimPlot(Tcells, "Tage", reduction.use = 'umap',
#                color.panel = dittoColors()[c(1:3,9:11)])
#   )

```

# Let's code up the cell highlight umaps.
## variables for highlight loop

```{r make some variables for cell highlight function}
#unique identities?



#colors corresponding with unique identities?


```

## set up highlight umap function + loop
```{r plot cell highlights, fig.height=3, fig.width=4}

# How can we turn this plot into a function?
dittoDimPlot(Tcells, "Tage", size = 0.5, reduction.use = "umap",
             cells.use = Tcells$Tage=="CD4_adult", colors = dittoColors()[1], 
             legend.show = FALSE, ylab = NULL, xlab = NULL, main = NULL,
             show.axes.numbers = FALSE)

# How can we use a for loop to generate plots for each of the variables we need?
# lets set up a for loop together!

for ( x in 1:6 ){
  
}

```

## add main plot and legend to plotlist list
```{r add main umap and legend to list of plots}
plotlist[[7]] <- main_umap
plotlist[[8]] <- legend
```

https://twitter.com/hadleywickham/status/643381054758363136

## final fancy combination!

```{r, combine figure 1A fig.height=6, fig.width=6}
plots = gridExtra::grid.arrange(
  grobs = plotlist,
  layout_matrix = matrix(c(
    7,7,7,2,
    7,7,7,6,
    7,7,7,4,
    1,5,3,8),
    ncol = 4))
```


# inspect some biology to set up next plots 1B:D
```{r res vs age comparison, fig.height=3, fig.width=4}
dittoDimPlot(Tcells, "Age", reduction.use = 'umap')
dittoDimPlot(Tcells, "RNA_snn_res.0.2", reduction.use = 'umap')
```

## set up new meta column
```{r make age/cluster vector}
age.clust <- factor(Tcells@meta.data$RNA_snn_res.0.2, 
                    labels = 
                      c("adult-cluster",
                        "UCB-cluster",
                        "fetal-cluster"))
#add this to metadata!

```

## plot umap 1B
```{r, fig.height=2, fig.width=4}
dittoBarPlot(
    object = Tcells, var = "age.cluster", group.by = "Sample")

#subset to CD4
#assign Y label
#remove legend
#relabel & rotate X axis labes
#remove main xlab
#fix colors
```
